---
title: "ODE"
author: "Wenmin Zhang"
date: "30/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

With the same initial condtion, we have that when $r=0.5$, the population eventually dies out and when $r=2.8$, the population acheives equilibrium after 10 generations while when $r=3.3$, the population cannot acheive equilibrium. 

A heuristic explanation will be that if the population achieves equilibrium than it can only be stable at $\frac{r-1}{r}$ or 0. Since the population size cannot go negative, thus 0 will be the stable point $r=0.5$ and 0.64 will be the stable point when $r=2.8$. The oscillation behavior at $r=3.3$ could due to that the change in the system is too rapid to reach a stable point.
```{r}

num=matrix(0,51,3)
num[1,]=rep(0.1,3)
rate = c(0.5,2.8,3.3)

for (c in 1:3)
  for (r in 1:50)
    num[r+1,c] = rate[c]*num[r,c]*(1-num[r,c])
    
```

```{r}
for (i in 1:3)
  plot(1:51,num[,i],type="b",col="red",xlab="time",ylab="number")
```

## Q2 

As the step size gets smaller, the estimate gets closer to the true value. This is expected since the appproximate procedure gets more accurate.
```{r}
integrate <- function(x0,t,h){
  x = x0
  ite = t/h
  for (i in 1:ite)
    x = x + h*x
    ite=ite+1
    return(x)
}
```

```{r}
integrate(1,5,0.1)
integrate(1,5,0.01)
integrate(1,5,0.001)
```

## Q3

(a)

```{r}
Cu=0.5
h=0.01
t=400
e=0.08
b=0.8
a=0.7
ite = t/h
v = rep(0,ite+1)
w = rep(0,ite+1)
v[1]=1
w[1]=0.1
```

```{r}
for (i in 1:ite) {
  dw = (e * (v[i] + a - b*w[i]))*h
  dv = (v[i] - (v[i]^3)/3 -w[i] + Cu)*h
  w[i+1] <- w[i] + dw
  v[i+1] <- v[i] + dv
#  cat(i,collapse="\n")
#  cat(dw,collapse="\n")
#  cat(dv,collapse="\n")
}
```

```{r}
plot(1:(ite+1),v,type="b",col="red",xlab="time",ylab="v")
```

(b)

```{r}
plot(w,v)
v0 <- seq(-2,2,by = 0.01)
w0 <- v0 - v0^3 / 3 + Cu
lines(w0,v0,col="red")
w1 <- (v0 + a)/b
lines(w1,v0,col="blue")
```

(c)

```{r}
#v - v^3/3 - (v + a)/b + Cu
polyroot(c(-a/b+Cu,1-1/b,0,-1/3))
```

(d)

The Jacobian matrix has the following entries accordingly: $1-v^2$, $-1$, $\epsilon$, $-b\epsilon$.

We can alternatively guess the sign of the eigenvalues from the spiral shape of the trajectory.

```{r}
vstar = polyroot(c(-a/b+Cu,1-1/b,0,-1/3))
eigen(matrix(c(1-vstar[2]^2,e,-1,-b*e),2,2))
```

(e)

```{r}
Cu = 0
for (i in 1:ite) {
  dw = (e * (v[i] + a - b*w[i]))*h
  dv = (v[i] - (v[i]^3)/3 -w[i] + Cu)*h
  w[i+1] <- w[i] + dw
  v[i+1] <- v[i] + dv
#  cat(i,collapse="\n")
#  cat(dw,collapse="\n")
#  cat(dv,collapse="\n")
}
plot(1:(ite+1),v,type="b",col="red",xlab="time",ylab="v")
plot(w,v)
v0 <- seq(-2,2,by = 0.01)
w0 <- v0 - v0^3 / 3 + Cu
lines(w0,v0,col="red")
w1 <- (v0 + a)/b
lines(w1,v0,col="blue")
polyroot(c(-a/b+Cu,1-1/b,0,-1/3))
vstar = polyroot(c(-a/b+Cu,1-1/b,0,-1/3))
eigen(matrix(c(1-vstar[2]^2,e,-1,-b*e),2,2))
```

(f)

```{r}
vs=c()
for (Cu in seq(0,0.5,0.01)){
  vs = c(vs,polyroot(c(-a/b+Cu,1-1/b,0,-1/3))[2])
}
```

```{r}
vseigen <- c()
for (vst in vs){
  tmp = eigen(matrix(c(1-vst^2,e,-1,-b*e),2,2))
#  cat(tmp$values)
  vseigen <- c(vseigen,Re(tmp$values[1]))
}
plot(seq(0,0.5,0.01),vs,col=ifelse(vseigen>0,"red","blue"))
```

